Community-Conditioned Language Models

### 2020-11-18 paper/paper.tex

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths.
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use the entropy metric
which is has stronger information-theoretic grounds, and is a measure
of the difficulty of identifying a community, as estimated by the
model. We call this measure the (linguistic) indiscernibility of the
community $c_j$, and use \(\Ind_j\) to stand for it in formulas.

Technically, it is defined as the exponential of the
entropy of the distribution \(D_j\), which is the (discrete)
probability distribution of predicted community for a random message from
$c_j$. $D_j$ is simply the $j$th row in \(C\). In sum, we have the formula:
\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

Indiscernibility is a robust measure: it is almost perfectly
correlated for various CCLM (transformer, lstm and various $c$) $M$.
\jp{other r's}
\begin{figure}
  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
  \caption{%
    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's r = \num{0.99} p < \num{0.0001}).
  }
  \label{fig:lmcc_ppl}
\end{figure}

However we observe that the CCLM perplexity is only weakly correlated with the
indiscernibility to the corresponding latent LMCC
(\cref{fig:cclm_lmcc_ppl}).
\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean linguistic perplexity (x-axis) and community indiscernibility
    (y-axis) per community (Pearson's r = \num{0.30}, p =
    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
    other models exhibit similar behavior.  }
  \label{fig:cclm_lmcc_ppl}
\end{figure}

>>>>

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths.
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(\cref{tab:model-ppls}).

@@

However, we want to measure the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use the entropy metric,
which is has stronger information-theoretic grounds, and is a measure
of the uncertainty of the LMCC, given messages from a particular 
community. 
Technically, the linguistic indiscernibility of community $c_j$ is defined 
as the exponential of the entropy of the $j$ row in \(C\),
which is a probability distribution 
of predicted community for a random message from $c_j$.
In sum, we have the formula: \bn{Took out $D$. is it needed elsewhere?}
\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]

Indiscernibility is a robust measure: it is almost perfectly
correlated for various CCLM (transformer, LSTM and various $c$) $M$.
In fact, we observe a correlation between the indiscernibility 
measures by every pair of models with Pearson's r $\geq 0.98$ 
and $p < 0.0001$ (see, for example, \cref{fig:lmcc_ppl}).

\begin{figure}
  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
  \caption{%
    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc_ppl}
\end{figure}

However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1. However the r is still very weakly positive for every model.}
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
\begin{figure}
  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
  \caption{%
    Mean community-wise linguistic perplexity (x-axis) 
    and indiscernibility (y-axis) 
    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
    Shown here for the LSTM with $c = 1$.}
  \label{fig:cclm_lmcc_ppl}
\end{figure}

\begin{table}
  \centering
  \input{floats/cclm_lmcc_ppl.tex}
  \caption{
  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
  }
  \label{tab:cclm_lmcc_ppl}
\end{table}
